### 2021-02-23 paper2/1-introduction.tex
Lots of improvements

Without due consideration for the data collection process, machine learning algorithms can exacerbate biases, or even introduce new ones if proper control is not exerted over their learning \cite{holstein2019improving}. 
While most of these issues can be solved by
% While efforts should be made to
controlling and curating data collection in a fairness-conscious fashion, 
% aiming to eliminate algorithmic bias during the early stages of the pipeline, 
doing so is not always an option, such as when working with historical data.
Efforts to address this problem algorithmically have been centred on developing statistical definitions of fairness and learning models that satisfy these definitions.
One popular definition of fairness used to guide the training of fair classifiers, for example, is \emph{demographic parity}, stating that positive outcome rates should be equalised (or \emph{invariant}) across protected groups.

In the typical setup, we have an input $\bm{x}$, a sensitive attribute $s$ that represents some non-admissible information like gender
and a class label $y$ which is the prediction target.
The idea of fair \emph{representation} learning \cite{ZemWuSwePitetal13}\cite{edwardsstorkey}\cite{madras2018learning}
is then to transform the input $\bm{x}$ to a representation $\bm{z}$ which is invariant to $s$.
Thus, learning from $\bm{z}$ will not introduce a forbidden dependence on $s$.
% $\bm{z}$ can then be used to learn a predictor
% In fair representation learning \cite{ZemWuSwePitetal13}\cite{edwardsstorkey}\cite{madras2018learning}, the goal is then to find a representation $\bm{z}$ that is invariant to $s$.
A good fair representation is one that preserves most of the information from $\bm{x}$ while satisfying the aforementioned constraints.
% While fair \emph{representations} are by no means the only method of achieving this goal, approaches that explicitly depend on the target label \cite{kamiran2012data} are restrictive in that they do not readily admit transfer to unseen tasks, and are bound by the requirement of having $s$ and $y$ labels be present together.

As unlabelled data is much more freely available than labelled data, it is of interest to learn the representation in an unsupervised manner.
This will allow us to draw on a much more diverse pool of data to learn from.
% ; this is particularly pertinent for fair representation learning as unlabelled data affords us a much more diverse pool of data to learn from.
%To this end, \cite{locatello2019fairness} recently made the connection from disentangled representations to fair representations:
%a model for disentanglement trained without knowledge of the sensitive attribute $s$
%nevertheless appears to improve fairness measures with respect to $s$.
%However, as pointed out by \cite{locatello2019challenging},
%some supervision or inductive bias is needed in order to recognise a well-disentangled representation.
While annotations for $y$ are often hard to come by (and often noisy \cite{KehCheQua18}),
annotations for the sensitive attribute $s$ are usually less so, as $s$ can often be obtained from demographic information provided by census data.
We thus consider the setting where the representation is learned from data
that is only labelled with $s$ and not $y$.
This is in contrast to most other representation learning methods.
% We thus consider the setting where data labelled with $s$ (partially labelled data) is available for learning the fair representation.
% Furthermore, we assume the existence of a set with $s$ labels
% whose distribution matches the deployment setting
We call the set used to learn the representation the \emph{representative} set,
because its distribution is meant to match the distribution of the deployment setting
(and is thus representative).
% We call this set the \emph{representative} set and its distribution is meant to match the distribution of the deployment setting.

Once we have learnt the mapping from $\bm{x}$ to $\bm{z}$,
we can transform the \emph{training} set which, in contrast to the representative set, has the $y$ labels (and $s$ labels).
In order to make our method more widely applicable,
we consider an \emph{aggravated fairness problem}
% we allow the case
in which the training set contains a strong spurious correlation between $s$ and $y$,
which makes it impossible to learn from it a representation which is invariant to $s$ but not invariant to $y$.
Non-invariance to $y$ is important in order to be able to predict $y$.
The training set thus does \emph{not} match the deployment setting,
thereby rendering the representative set essential for learning the right invariance.
% We also tackle the related problem of learning from biased data, specifically cases in which the training set (with $y$ labels) contains a strong spurious correlation between $s$ and $y$, and thus does not match the deployment setting.
From hereon, we will use the terms \emph{spurious} and \emph{sensitive} interchangeably, depending on the context, to refer to an attribute of the data we seek invariance to.
% This is essentially a form of strong sampling bias and is not an unrealistic complication, having been shown to plague real-world datasets \cite{kallus2018residual}.
%Classifiers trained on ImageNet for example, have been shown to be biased towards texture \cite{Geir18}. \cite{zhang2018visual} similarly examine the problem of learning from biased data, showing that pre-trained models can exploit biases in the data by learning patterns semantically unrelated to the target, an issue that can be difficult to identify when the bias pervades both the training and test sets.
We can draw a connection between learning in the presence of spurious correlations and what \cite{kallus2018residual} call \emph{residual unfairness}.
Consider the Stop, Question and Frisk (SQF) dataset for example:
the data was collected in New York City, but the demographics of the recorded cases do not represent the true demographics of NYC well.
The demographic attributes of the recorded individuals might correlate so strongly with the prediction target that the two are nearly indistinguishable.
This is the scenario that we are investigating: $s$ and $y$ are so closely correlated in the labelled dataset that they cannot be distinguished, but the learning of $s$ is favoured due to being the ``path of least resistance''.
The deployment setting (i.e.\ the test set) does not possess this strong correlation and thus a na\"ive approach will lead to very unfair predictions.
In this case, a disentangled representation is insufficient;
the representation needs to be explicitly invariant solely with respect to $s$.
In our approach, we make use of the (partially labelled) representative set to learn this invariant representation.

While there is a substantial body of literature devoted to the problems of fair representation-learning, exactly how the invariance in question is achieved is often overlooked. 
When critical decisions, such as who should receive bail or be released from jail, are being deferred to an automated decision making system, it is critical that people be able to trust the logic of the model underlying it, whether it be via semantic or visual explanations. 
We build on the work of \cite{QuaShaTho19} and learn a decomposition ($f^{-1}: Z_s \times Z_{\neg s} \rightarrow X$) of the \emph{data domain} ($X$) into independent subspaces \emph{invariant} to  $s$ ($Z_{\neg s}$) and \emph{indicative} of $s$ ($Z_{s}$), which lends an interpretability that is absent from most representation-learning methods.
While model interpretability has no strict definition \cite{zhang2018visual}, we follow the intuition of \cite{adel2018discovering} -- \emph{a simple relationship to something we can understand}, a definition which representations in the data domain naturally fulfil.

@@

To render our representations interpretable, we rely on a simple transformation we call \emph{null-sampling} to map invariant representations in the data domain.
Previous approaches to fair representation learning \cite{beutel,edwardsstorkey,madras2018learning,LouSweLi15} predominantly rely upon autoencoder models to jointly minimise reconstruction loss and invariance.
We discuss first how this can be done with such a model that we refer to as cVAE (conditional VAE),
before arguing that the bijectivity of invertible neural networks (INNs)~\cite{Dinh2014} makes them better suited to this task. 
We refer to the variant of our method based on these as cFlow (conditional Flow).
INNs have several properties that make them appealing for unsupervised representation learning. 
The focus of our approach is on creating invariant representations that preserve the non-sensitive information maximally, with only knowledge of $s$ and not of the target $y$, while at the same time having the ability to easily probe what has been learnt.

>>>>

Without due consideration for the data collection process, machine learning algorithms can exacerbate biases, or even introduce new ones if proper control is not exerted over their learning \citep{holstein2019improving}. 
While most of these issues can be solved by
% While efforts should be made to
controlling and curating data collection in a fairness-conscious fashion, 
% aiming to eliminate algorithmic bias during the early stages of the pipeline, 
doing so is not always an option, such as when working with historical data.
Efforts to address this problem algorithmically have been centred on developing statistical definitions of fairness and learning models that satisfy these definitions.
One popular definition of fairness used to guide the training of fair classifiers, for example, is \emph{demographic parity}, stating that positive outcome rates should be equalised (or \emph{invariant}) across protected groups.

In the typical setup, we have an input $\bm{x}$, a sensitive attribute $s$ that represents some non-admissible information like gender
and a class label $y$ which is the prediction target.
The idea of fair \emph{representation} learning \citep{ZemWuSwePitetal13,edwards2016censoring,madras2018learning}
is then to transform the input $\bm{x}$ to a representation $\bm{z}$ which is invariant to $s$.
Thus, learning from $\bm{z}$ will not introduce a forbidden dependence on $s$.
% $\bm{z}$ can then be used to learn a predictor
% In fair representation learning \cite{ZemWuSwePitetal13}\cite{edwards2016censoring}\cite{madras2018learning}, the goal is then to find a representation $\bm{z}$ that is invariant to $s$.
A good fair representation is one that preserves most of the information from $\bm{x}$ while satisfying the aforementioned constraints.
% While fair \emph{representations} are by no means the only method of achieving this goal, approaches that explicitly depend on the target label \cite{kamiran2012data} are restrictive in that they do not readily admit transfer to unseen tasks, and are bound by the requirement of having $s$ and $y$ labels be present together.

As unlabelled data is much more freely available than labelled data, it is of interest to learn the representation in an unsupervised manner.
This will allow us to draw on a much more diverse pool of data to learn from.
% ; this is particularly pertinent for fair representation learning as unlabelled data affords us a much more diverse pool of data to learn from.
%To this end, \cite{locatello2019fairness} recently made the connection from disentangled representations to fair representations:
%a model for disentanglement trained without knowledge of the sensitive attribute $s$
%nevertheless appears to improve fairness measures with respect to $s$.
%However, as pointed out by \cite{locatello2019challenging},
%some supervision or inductive bias is needed in order to recognise a well-disentangled representation.
While annotations for $y$ are often hard to come by (and often noisy; see~\cite{kehrenberg2020tuning}),
annotations for the sensitive attribute $s$ are usually less so, as $s$ can often be obtained from demographic information provided by census data.
We thus consider the setting where the representation is learned from data
that is only labelled with $s$ and not $y$.
This is in contrast to most other representation learning methods.
% We thus consider the setting where data labelled with $s$ (partially labelled data) is available for learning the fair representation.
% Furthermore, we assume the existence of a set with $s$ labels
% whose distribution matches the deployment setting
We call the set used to learn the representation the \emph{representative} set,
because its distribution is meant to match the distribution of the deployment setting
(and is thus representative).
% We call this set the \emph{representative} set and its distribution is meant to match the distribution of the deployment setting.

Once we have learnt the mapping from $\bm{x}$ to $\bm{z}$,
we can transform the \emph{training} set which, in contrast to the representative set, has the $y$ labels (and $s$ labels).
In order to make our method more widely applicable,
we consider an \emph{aggravated fairness problem}
% we allow the case
in which the training set contains a strong spurious correlation between $s$ and $y$,
which makes it impossible to learn from it a representation which is invariant to $s$ but not invariant to $y$.
Non-invariance to $y$ is important in order to be able to predict $y$.
The training set thus does \emph{not} match the deployment setting,
thereby rendering the representative set essential for learning the right invariance.
% We also tackle the related problem of learning from biased data, specifically cases in which the training set (with $y$ labels) contains a strong spurious correlation between $s$ and $y$, and thus does not match the deployment setting.
From hereon, we will use the terms \emph{spurious} and \emph{sensitive} interchangeably, depending on the context, to refer to an attribute of the data we seek invariance to.
% This is essentially a form of strong sampling bias and is not an unrealistic complication, having been shown to plague real-world datasets \cite{kallus2018residual}.
%Classifiers trained on ImageNet for example, have been shown to be biased towards texture \cite{Geir18}. \cite{zhang2018visual} similarly examine the problem of learning from biased data, showing that pre-trained models can exploit biases in the data by learning patterns semantically unrelated to the target, an issue that can be difficult to identify when the bias pervades both the training and test sets.
We can draw a connection between learning in the presence of spurious correlations and what \citet{kallus2018residual} call \emph{residual unfairness}.
Consider the Stop, Question and Frisk (SQF) dataset for example:
the data was collected in New York City, but the demographics of the recorded cases do not represent the true demographics of NYC well.
The demographic attributes of the recorded individuals might correlate so strongly with the prediction target that the two are nearly indistinguishable.
This is the scenario that we are investigating: $s$ and $y$ are so closely correlated in the labelled dataset that they cannot be distinguished, but the learning of $s$ is favoured due to being the ``path of least resistance''.
The deployment setting (i.e.\ the test set) does not possess this strong correlation and thus a na\"ive approach will lead to very unfair predictions.
In this case, a disentangled representation is insufficient;
the representation needs to be explicitly invariant solely with respect to $s$.
In our approach, we make use of the (partially labelled) representative set to learn this invariant representation.

While there is a substantial body of literature devoted to the problems of fair representation-learning, exactly how the invariance in question is achieved is often overlooked. 
When critical decisions, such as who should receive bail or be released from jail, are being deferred to an automated decision making system, it is critical that people be able to trust the logic of the model underlying it, whether it be via semantic or visual explanations. 
We build on the work of \citet{QuaShaTho19} and learn a decomposition ($f^{-1}: Z_s \times Z_{\neg s} \rightarrow X$) of the \emph{data domain} ($X$) into independent subspaces \emph{invariant} to  $s$ ($Z_{\neg s}$) and \emph{indicative} of $s$ ($Z_{s}$), which lends an interpretability that is absent from most representation-learning methods.
While model interpretability has no strict definition \citep{zhang2018visual}, we follow the intuition of \citet{adel2018discovering} -- \emph{a simple relationship to something we can understand}, a definition which representations in the data domain naturally fulfil.

@@

To render our representations interpretable, we rely on a simple transformation we call \emph{null-sampling} to map invariant representations in the data domain.
Previous approaches to fair representation learning \citep{beutel2017data,edwards2016censoring,madras2018learning,louizos2016variational} predominantly rely upon autoencoder models to jointly minimise reconstruction loss and invariance.
We discuss first how this can be done with such a model that we refer to as cVAE (conditional VAE),
before arguing that the bijectivity of invertible neural networks (INNs)~\citep{Dinh2014} makes them better suited to this task. 
We refer to the variant of our method based on these as cFlow (conditional Flow).
INNs have several properties that make them appealing for unsupervised representation learning. 
The focus of our approach is on creating invariant representations that preserve the non-sensitive information maximally, with only knowledge of $s$ and not of the target $y$, while at the same time having the ability to easily probe what has been learnt.

### 2021-02-23 paper2/2-related-work.tex
Some meager progress
Lots of improvements

% \subsubsection{Learning fair representations.}
\textbf{Learning fair representations.}
%As we have alluded to, the goal of producing invariant representations is similar to that of producing \emph{fair} representations.
% In fairness problems, there is usually a \emph{sensitive attribute}, $s$ (for example, gender or race),
% that should not be used to make decisions.
Given a sensitive attribute $s$ (for example, gender or race) and inputs $\bm{x}$,
a fair representation $\bm{z}$ of $\bm{x}$ is then one for which $\bm{z} \perp s$ holds,
while ideally also being predictive of the class label $y$.
\cite{ZemWuSwePitetal13} was the first to propose the learning of fair representations which allow for transfer to new classification tasks.
More recent methods are often based on variational autoencoders (VAEs)~\cite{kingma2013auto,LouSweLi15,edwardsstorkey,beutel}.
The achieved fairness of the representation can be measured with various fairness metrics.
These measure, however, usually how fair the predictions of a classifier are
and not how fair a representation is.
% defined with respect to predictions and not representations.

% There is not one metric universally best-suited to measuring the invariance (or fairness) of representations.
The appropriate measure of fairness for a given task is domain-specific \cite{liu2018delayed}
and there is often not a universally accepted measure.
However, \emph{Demographic Parity} is the most widely used~\cite{LouSweLi15,edwardsstorkey,beutel}.
Demographic Parity demands $\hat{y} \perp s$ where $\hat{y}$ refers to the predictions of the classifier.
In the context of fair representations, we measure the Demographic Parity of a downstream classifier, $f(\cdot )$, which is trained on the representation $z$ i.e.  $f: Z \to \hat{Y}$.

@@

The majority of existing methods for fair representations also make use of $y$ labels during training,
in order to ensure that $\bm{z}$ remains predictive of $y$.
This aspect can, in theory, be removed from the methods,
but then there is no guarantee that information about $y$ is preserved \cite{LouSweLi15}. 
% Existing methods designed to create fair representations can, in theory, be extended to the regime in which only the $s$, and not the $y$, labels are available. 
% However, it is not without its drawbacks as, in removing $s$, there is no guarantee that information about $y$ is preserved \cite{LouSweLi15}. 
% For this reason, $y$ is typically supplied during training and the representation encouraged to be predictive of it.

%\subsection{Learning Fair and Transferable Representations}
% \subsubsection{Learning fair representations that are transferable and in the data domain.}
\textbf{Learning fair, transferrable representations.}
% Outside of computer vision, \cite{madras2018learning} have also worked on removing a problematic spurious correlation.
In addition to producing fair representations, \cite{madras2018learning} want to ensure the representations are transferrable.
Here, an adversary is used to remove sensitive information from a representation $z$.
Auxiliary prediction and reconstruction networks, to predict class label $y$
% to ensure it remains predictive of $y$
and reconstruct the input $x$ respectively,
are trained on top of $z$, with $s$ being ancillary input to the reconstruction.
% alongside a reconstruction loss computed with respect to the output of a decoder that takes $z$ and the sensitive label $s$ as input is added. 
% This is so that $x$ can still be reconstructed from $z$, despite the removal of $s$.
%The decoder is utilised only for the purpose of maximum likelihood learning of the data distribution.
% By conditioning the decoder on the sensitive attribute, information about it can be ``off-loaded'' from the encoder such that information about $s$ need not be contained in the fair representation while still permitting the use of a reconstruction loss needed to capture non-sensitive information.
% The decoder plays a role only in the loss function.
% In contrast, we make explicit use of the decoder not only for characterising the behaviour of the model and also for evaluation.

Also related is \cite{creager2019flexibly} who employ a FactorVAE \cite{kim2018disentangling} regularised for fairness.
The idea is to learn a representation that is both disentangled and invariant to multiple sensitive attributes.
This factorisation makes the latent space easily manipulable such that the different subspaces can be freely removed and composed at test time.
Zeroing out the dimensions or replacing them with independent noise imparts invariance to the corresponding sensitive attribute.
This method closely resembles ours when we use an invertible encoder.
However, the emphasis of our approach is on interpretability, information-preservation, and coping with sampling bias - especially extreme cases where $|\, \textrm{supp}(S_{tr} \times Y_{tr}) | < |\, \textrm{supp}(S_{te} \times Y_{te}) |$.
% Namely, the invertibility of the network means we can optimise for invariance singularly without the burden of a reconstruction loss.
% While we do not explicitly consider the case of multi-attribute fairness, our method can be easily adapted for this use-case.

Attempts were made by \cite{QuaShaTho19} prior to this work to learn fair representations in the data domain in order to make it interpretable and transferable.
In their work, the input is assumed to be additively decomposable in the feature space into a \emph{fair} and \emph{unfair} component, which together can be used by the decoder to recover the original input.
This allows us to examine representations in a human-interpretable space and confirm that the model is not learning a relationship reliant on a sensitive attribute.
Though a first step in this direction, we believe such a linear decomposition is not sufficiently expressive to fully capture the relationship between the sensitive and non-sensitive attributes.
Our approach allows for the modelling of more complex relationships.

% \subsubsection{Learning in the presence of spurious correlations.}
\textbf{Learning in the presence of spurious correlations.}
% As  previously  discussed,  the  goal  of  producing  fair representations is similar to the goal of producing representations invariant to spurious correlations found in the training data.
Strong spurious correlations make the task of learning a robust classifier challenging: the classifier may learn to exploit correlations unrelated to the true causal relationship between the features and label, and thereby fail to generalise to novel settings.
This problem was recently tackled by \cite{ln2l} who apply a penalty based on the mutual information between the feature embedding and the spurious variable. 
While the method is effective under mild biasing, we show experimentally that it is not robust to the range of settings we consider.

Jacobsen et al. \cite{JacBehZemBet19} explore the vulnerability of traditional neural networks to spurious variables -- e.g., textures, in the case of ImageNet \cite{Geir18} -- and propose a INN-based solution akin to ours.
The INN's encoding is split such that one partition, $z_b$ is encouraged to be predictive of the spurious variable while the other serves as the logits for classification of the semantic label. 
Information related to the nuisance variable is ``pulled out'' of the logits as a result of maximising $\log p(s|z_n)$.
This specific approach, however, is incompatible with the settings we consider, due to its requirement that both $s$ and $y$ be available at training time.

Viewing the problem from a causal perspective, \cite{arjovsky2019invariant} develop a variant of empirical risk minimisation called invariant risk minimisation (IRM).
The goal of IRM is to train a predictor that generalises across a large set of unseen environments; because variables with spurious correlations do not represent a stable causal mechanism, the predictor learns to be invariant to them. IRM assumes that the training data is not \emph{iid} but is partitioned into distinct environments, $e \in E$. The optimal predictor is then defined as the minimiser of the sum of the empirical risk $R_e$ over this set. In contrast, we assume possession of only a single source of \emph{labelled}, albeit spuriously-correlated, data, but that we have a second source of data that is free of spurious correlations, with the benefit being that it only needs to be labelled \emph{with respect to $s$}.

% The model thereby enforces their independence.
% This is achieved with an adversarial approach, borrowing the gradient reversal technique from \cite{ganin2016domain}.
%The authors construct the coloured MNIST dataset in two steps.
%First, ten distinct colours are assigned to each digit uniquely; these colours parameterise the means of ten corresponding Gaussian distributions from which colour samples are drawn.
%The standard deviation ($\sigma$) of the Gaussian distribution controls the dispersion of the sampled colours around these means.
% To demonstrate the effectiveness of their model, \cite{ln2l} construct a coloured version of the MNIST dataset as follows.
% During training, colours are sampled from a Gaussian distribution (with standard deviation $\sigma$) where each digit is associated with a single fixed mean colour.
%the colours are sampled abiding by this one-to-one colour mapping;
% At test time however, a colour mean is chosen at random from the 10 mean colours used during training.
% The actual colour is sampled with the same $\sigma$ as in the training set.
%there is no such designation and colours are sampled randomly and unrestrictedly from the complete palette.
%As such, a classifier that lazily minimises its loss by treating the pixel values as a lookup table falls flat at inference owing to a shift in the distribution of the spurious variables away from that of the target's.
% We follow this approach to evaluate performance of our NoSiNN framework in a synthetic setting (see Fig. \ref{fig:cmnist} for qualitative results).

% For the training strategy of the \cite{ln2l} model, a neural network is trained to predict the digit class, $y$, while an adversarial network takes one of the intermediate layers as input to predict the spurious value, colour.
% The first network seeks to prevent the adversary from making correct predictions, which means discarding or obfuscating information about colour.
% For this approach to work, the adversary needs to be able to distinguish between the digit class and the colour.
% To do this, the adversary is allowed access to the sampled RGB values of the colour that it is trying to remove, and not just the mean.
% As the sampled colour varies according to the standard deviation of the Gaussian distribution, the actual colour and the digit class vary in correlation.
% As the colour becomes less descriptive of the digit class, the network learns to disentangle the two.
% This works better, the larger $\sigma$; a major limitation of the approach its failure to deal with extremely low $\sigma$ values.

>>>>

\subsection{Learning fair representations.}
%As we have alluded to, the goal of producing invariant representations is similar to that of producing \emph{fair} representations.
% In fairness problems, there is usually a \emph{sensitive attribute}, $s$ (for example, gender or race),
% that should not be used to make decisions.
Given a sensitive attribute $s$ (for example, gender or race) and inputs $\bm{x}$,
a fair representation $\bm{z}$ of $\bm{x}$ is then one for which $\bm{z} \perp s$ holds,
while ideally also being predictive of the class label $y$.
\citet{ZemWuSwePitetal13} was the first to propose the learning of fair representations which allow for transfer to new classification tasks.
More recent methods are often based on variational autoencoders (VAEs)~\citep{kingma2013auto,louizos2016variational,edwards2016censoring,beutel2017data}.
The achieved fairness of the representation can be measured with various fairness metrics.
These measure, however, usually how fair the predictions of a classifier are
and not how fair a representation is.
% defined with respect to predictions and not representations.

% There is not one metric universally best-suited to measuring the invariance (or fairness) of representations.
The appropriate measure of fairness for a given task is domain-specific \citep{liu2018delayed}
and there is often not a universally accepted measure.
However, \emph{Demographic Parity} is the most widely used~\citep{louizos2016variational,edwards2016censoring,beutel2017data}.
Demographic Parity demands $\hat{y} \perp s$ where $\hat{y}$ refers to the predictions of the classifier.
In the context of fair representations, we measure the Demographic Parity of a downstream classifier, $f(\cdot )$, which is trained on the representation $z$ i.e.  $f: Z \to \hat{Y}$.

@@

The majority of existing methods for fair representations also make use of $y$ labels during training,
in order to ensure that $\bm{z}$ remains predictive of $y$.
This aspect can, in theory, be removed from the methods,
but then there is no guarantee that information about $y$ is preserved \citep{louizos2016variational}. 
% Existing methods designed to create fair representations can, in theory, be extended to the regime in which only the $s$, and not the $y$, labels are available. 
% However, it is not without its drawbacks as, in removing $s$, there is no guarantee that information about $y$ is preserved \cite{louizos2016variational}. 
% For this reason, $y$ is typically supplied during training and the representation encouraged to be predictive of it.

\subsection{Learning fair, transferrable representations}
% Outside of computer vision, \cite{madras2018learning} have also worked on removing a problematic spurious correlation.
In addition to producing fair representations, \citet{madras2018learning} want to ensure the representations are transferrable.
Here, an adversary is used to remove sensitive information from a representation $z$.
Auxiliary prediction and reconstruction networks, to predict class label $y$
% to ensure it remains predictive of $y$
and reconstruct the input $x$ respectively,
are trained on top of $z$, with $s$ being ancillary input to the reconstruction.
% alongside a reconstruction loss computed with respect to the output of a decoder that takes $z$ and the sensitive label $s$ as input is added. 
% This is so that $x$ can still be reconstructed from $z$, despite the removal of $s$.
%The decoder is utilised only for the purpose of maximum likelihood learning of the data distribution.
% By conditioning the decoder on the sensitive attribute, information about it can be ``off-loaded'' from the encoder such that information about $s$ need not be contained in the fair representation while still permitting the use of a reconstruction loss needed to capture non-sensitive information.
% The decoder plays a role only in the loss function.
% In contrast, we make explicit use of the decoder not only for characterising the behaviour of the model and also for evaluation.

Also related is \citet{creager2019flexibly} who employ a FactorVAE \citep{kim2018disentangling} regularised for fairness.
The idea is to learn a representation that is both disentangled and invariant to multiple sensitive attributes.
This factorisation makes the latent space easily manipulable such that the different subspaces can be freely removed and composed at test time.
Zeroing out the dimensions or replacing them with independent noise imparts invariance to the corresponding sensitive attribute.
This method closely resembles ours when we use an invertible encoder.
However, the emphasis of our approach is on interpretability, information-preservation, and coping with sampling bias - especially extreme cases where $|\, \textrm{supp}(S_{tr} \times Y_{tr}) | < |\, \textrm{supp}(S_{te} \times Y_{te}) |$.
% Namely, the invertibility of the network means we can optimise for invariance singularly without the burden of a reconstruction loss.
% While we do not explicitly consider the case of multi-attribute fairness, our method can be easily adapted for this use-case.

Attempts were made by~\citet{QuaShaTho19} prior to this work to learn fair representations in the data domain in order to make it interpretable and transferable.
In their work, the input is assumed to be additively decomposable in the feature space into a \emph{fair} and \emph{unfair} component, which together can be used by the decoder to recover the original input.
This allows us to examine representations in a human-interpretable space and confirm that the model is not learning a relationship reliant on a sensitive attribute.
Though a first step in this direction, we believe such a linear decomposition is not sufficiently expressive to fully capture the relationship between the sensitive and non-sensitive attributes.
Our approach allows for the modelling of more complex relationships.

\subsection{Learning in the presence of spurious correlations}
% As  previously  discussed,  the  goal  of  producing  fair representations is similar to the goal of producing representations invariant to spurious correlations found in the training data.
Strong spurious correlations make the task of learning a robust classifier challenging: the classifier may learn to exploit correlations unrelated to the true causal relationship between the features and label, and thereby fail to generalise to novel settings.
This problem was recently tackled by \citet{kim2019learning} who apply a penalty based on the mutual information between the feature embedding and the spurious variable. 
While the method is effective under mild biasing, we show experimentally that it is not robust to the range of settings we consider.

\citet{JacBehZemBet19} explore the vulnerability of traditional neural networks to spurious variables -- e.g., textures, in the case of ImageNet \citep{Geir18} -- and propose a INN-based solution akin to ours.
The INN's encoding is split such that one partition, $z_b$ is encouraged to be predictive of the spurious variable while the other serves as the logits for classification of the semantic label. 
Information related to the nuisance variable is ``pulled out'' of the logits as a result of maximising $\log p(s|z_n)$.
This specific approach, however, is incompatible with the settings we consider, due to its requirement that both $s$ and $y$ be available at training time.

Viewing the problem from a causal perspective, \citet{arjovsky2019invariant} develop a variant of empirical risk minimisation called invariant risk minimisation (IRM).
The goal of IRM is to train a predictor that generalises across a large set of unseen environments; because variables with spurious correlations do not represent a stable causal mechanism, the predictor learns to be invariant to them. IRM assumes that the training data is not \emph{iid} but is partitioned into distinct environments, $e \in E$. The optimal predictor is then defined as the minimiser of the sum of the empirical risk $R_e$ over this set. In contrast, we assume possession of only a single source of \emph{labelled}, albeit spuriously-correlated, data, but that we have a second source of data that is free of spurious correlations, with the benefit being that it only needs to be labelled \emph{with respect to $s$}.

% The model thereby enforces their independence.
% This is achieved with an adversarial approach, borrowing the gradient reversal technique from \cite{ganin2016domain}.
%The authors construct the coloured MNIST dataset in two steps.
%First, ten distinct colours are assigned to each digit uniquely; these colours parameterise the means of ten corresponding Gaussian distributions from which colour samples are drawn.
%The standard deviation ($\sigma$) of the Gaussian distribution controls the dispersion of the sampled colours around these means.
% To demonstrate the effectiveness of their model, \cite{kim2019learning} construct a coloured version of the MNIST dataset as follows.
% During training, colours are sampled from a Gaussian distribution (with standard deviation $\sigma$) where each digit is associated with a single fixed mean colour.
%the colours are sampled abiding by this one-to-one colour mapping;
% At test time however, a colour mean is chosen at random from the 10 mean colours used during training.
% The actual colour is sampled with the same $\sigma$ as in the training set.
%there is no such designation and colours are sampled randomly and unrestrictedly from the complete palette.
%As such, a classifier that lazily minimises its loss by treating the pixel values as a lookup table falls flat at inference owing to a shift in the distribution of the spurious variables away from that of the target's.
% We follow this approach to evaluate performance of our NoSiNN framework in a synthetic setting (see Fig. \ref{fig:cmnist} for qualitative results).

% For the training strategy of the \cite{kim2019learning} model, a neural network is trained to predict the digit class, $y$, while an adversarial network takes one of the intermediate layers as input to predict the spurious value, colour.
% The first network seeks to prevent the adversary from making correct predictions, which means discarding or obfuscating information about colour.
% For this approach to work, the adversary needs to be able to distinguish between the digit class and the colour.
% To do this, the adversary is allowed access to the sampled RGB values of the colour that it is trying to remove, and not just the mean.
% As the sampled colour varies according to the standard deviation of the Gaussian distribution, the actual colour and the digit class vary in correlation.
% As the colour becomes less descriptive of the digit class, the network learns to disentangle the two.
% This works better, the larger $\sigma$; a major limitation of the approach its failure to deal with extremely low $\sigma$ values.
